---
title: "Project2"
author: "Yuzhi Li, Katelyn Settlemyre"
date: "2022-10-03"
urlcolor: blue
output: 
  html_document:
    toc: TRUE
    toc_depth: 2
---

```{r packages, include=FALSE}
#packages needed to get API data and work with it
library(httr)
library(jsonlite)
library(tidyverse)
```
This vignette explains how to contact an API, demonstrated with the [Spoonacular Food API](https://spoonacular.com/food-api). We will create functions to interact with the API and return data. 

# Example of R-code that can contact Food data API by searching criteria

```{r}
res<-GET("https://api.spoonacular.com/recipes/complexSearch?query=chicken&maxFat=60&number=5&apiKey=be3c2a74ce6c4f0c83b254b34ceca7f5")
```

# Package Requirements

Here, we list the packages required to interact with the API and use the functions in this vignette.
- `httr` \- used to access the API  
- `jsonlite` \- used to convert the data from the API into a data frame  
- `tidyverse` \- used for data manipulation and graphing

# API Interaction Functions
To use some of the interaction functions in this section, you will need to have a key to access the API. You can register for free [here](https://spoonacular.com/food-api/console#Dashboard) if you do not already have a key for this API.

## `recipe.search`
This function allows you to search for recipes given query (i.e., what are you looking for?), [diet type](https://spoonacular.com/food-api/docs#Diets), [type of dish](https://spoonacular.com/food-api/docs#Meal-Types) (such as breakfast, dessert, side dish, etc.), the maximum number of carbs, the minimum amount of protein, and the maximum amount of fat. To give more than one option, you will need to give a string of comma separated values.
```{r}
recipe.search<- function(key=NULL, query=NULL, diet=NULL, type=NULL, maxCarbs=NULL, minProtein=NULL, maxFat=NULL){
  if(is.null(key)){
    return("your API Key is required.")
  }
  basic.url<-"https://api.spoonacular.com/recipes/complexSearch"
  search.opt<-list(query=query, diet=diet, type=type, maxCarbs=maxCarbs, minProtein=minProtein, maxFat=maxFat, apiKey=key)
  search.opt2<-search.opt[-which(sapply(search.opt, is.null))]
  res<-paste0(basic.url,"?")
  for(i in 1:length(search.opt2)){
    res<-paste0(res,"&", names(search.opt2[i]), "=", search.opt2[i])
  }
  data<-httr::GET(res)
  mydf<- jsonlite::fromJSON(data$url)
  result<-as.data.frame(mydf)
      return(result)
    }
```

Search recipe example by using the function above:

```{r}
recipe.search(key="be3c2a74ce6c4f0c83b254b34ceca7f5", query="chicken", maxFat=60)
```

## `random.recipes`
With this function, you can get between 1 and 100 random recipes. ~~You may also include tags by giving a string of comma separated values.~~
```{r}
# Gives full recipes, still editing so not included yet
#gives 1-100 random recipes
random.recipes <- function(apiKey=NULL, number=NULL, tags=NULL){
  if(is.null(apiKey)){
    return("Please enter your API key.")
  }
  
  baseURL <- "https://api.spoonacular.com/recipes/random"
  input <- list(apiKey=apiKey, number=number, tags=tags)
  opts <- input[-which(sapply(input, is.null))]
  link <- paste0(baseURL, "?")
  for(i in 1:length(opts)){
    link<-paste0(link,"&", names(opts[i]), "=", opts[i])
  }
  link <- str_remove(link, "&")    #remove the first &
  dat <- httr::GET(link)
  df<- jsonlite::fromJSON(dat$url)
  results <- as.data.frame(df)
  return(results)
}
```

## `diet.restrictions`
For people with dietary restrictions, intolerances, or sensitivities, this function allows you to specify [diet](https://spoonacular.com/food-api/docs#Diets) and food [intolerances](https://spoonacular.com/food-api/docs#Intolerances). ~~It also allows for ingredients to be excluded. To give more than one option, you will need to give a string of comma separated values.~~
```{r}
#function to search for recipes given diet type, food intolerances, excluded ingredients
diet.restrictions <- function(apiKey=NULL, diet=NULL, intolerances=NULL, exclude=NULL){
  if(is.null(apiKey)){
    return("Please enter your API key.")
  }
  
  baseURL <- "https://api.spoonacular.com/recipes/complexSearch"
  input <- list(diet = diet, intolerances = intolerances, excludeIngredients = exclude)
  opts <- input[-which(sapply(input, is.null))]
  link <- paste0(baseURL, "?apiKey=", apiKey)
  for(i in 1:length(opts)){
    link<-paste0(link,"&", names(opts[i]), "=", opts[i])
  }
  dat <- httr::GET(link)
  df<- jsonlite::fromJSON(dat$url)
  results <- as.data.frame(df)
  return(results)
}
```



# Data Exploration

We would like to compare two different diets of chicken recipes, one is for "ketogenic" and the other is for "whole30". 

The definitions are as below:  
* **Ketogenic**: The keto diet is based more on the ratio of fat, protein, and carbs in the diet rather than specific ingredients. Generally speaking, high fat, protein-rich foods are acceptable and high carbohydrate foods are not. The formula we use is 55-80% fat content, 15-35% protein content, and under 10% of carbohydrates.

* **Whole30**: Allowed ingredients include meat, fish/seafood, eggs, vegetables, fresh fruit, coconut oil, olive oil, small amounts of dried fruit and nuts/seeds. Ingredients not allowed include added sweeteners (natural and artificial, except small amounts of fruit juice), dairy (except clarified butter or ghee), alcohol, grains, legumes (except green beans, sugar snap peas, and snow peas), and food additives, such as carrageenan, MSG, and sulfites.

```{r}
recipe1<- recipe.search(key="be3c2a74ce6c4f0c83b254b34ceca7f5", query="chicken", diet="ketogenic", minProtein=5, maxCarbs=100)
recipe1

recipe2<- recipe.search(key="be3c2a74ce6c4f0c83b254b34ceca7f5", query="chicken", diet="whole30", minProtein=5, maxCarbs=100)
recipe2
```

We will compare the carbohydrates and protein amount in those two different diet groups.

* nutrition table for recipe1

```{r}
nutri1<-recipe1$results.nutrition
nutri1<-data.frame(t(matrix(unlist(nutri1),nrow=6))) %>% select(X3, X4) %>% rename(Protein=X3, Carbohydrates=X4) %>% add_column(Diet="keto")
nutri1
```

* nutrition table for recipe2

```{r}
nutri2<-recipe2$results.nutrition
nutri2<-data.frame(t(matrix(unlist(nutri2),nrow=6))) %>% select(X3, X4) %>% rename(Protein=X3, Carbohydrates=X4) %>% add_column(Diet="whole")
nutri2
```
Combine the above nutrition tables together.

```{r}
nutri<-bind_rows(nutri1, nutri2)
nutri
```


Looking at random recipes, we can specify how many recipes we want, and we can see how many fall into specific types. For example,how many of our recipes are vegetarian only, dairy free only, both, or neither.
```{r}
rand <- random.recipes(apiKey = "31bf40952ec14026b2f82b5055fdab93", number=20)
table(Vegetarian=rand$recipes.vegetarian, DairyFree=rand$recipes.dairyFree)
```


# Wrap-Up